---
title: "ISLR"
author: "GT"
date: "15 septembre 2016"
output: html_document
---

## Chapter 2

Load the `Auto` data set. This data set is part of the `ISLR` library.

```{r}
library(ISLR)
```


```{r}
dim(Auto)

# Check the variables names
names(Auto)
```

We can use the `plot()` function to produce _scatterplots_ of quantitative
variables.

```{r}
plot(Auto$cylinders, Auto$mpg)
```

The `cylinders` variable is stored as a numeric vector, so R has 
treated it as quantitative. However, since there are only a small
number possible values for cylinders, one may prefer to treat it
as a qualitative variable.

```{r}
# The as.factor() function converts quantitative variable into
# qualitative variable
Auto$cylinders = as.factor(Auto$cylinders)
```

If the variable plotted on the x-axis is categorical, then
_boxplots_ will be automatically be produced by the `plot()`
function.

```{r}
plot(Auto$cylinders, Auto$mpg, col="red", varwidth=T, xlab="cylinders",
     ylab="MPG")
```

The `hist()` function can be used to plot a _histogram_.

```{r}
hist(Auto$mpg, col=2, breaks=15)
```

The `pairs() function creates a _scatterplot matrix_.

```{r}
pairs(~ mpg + displacement + horsepower + weight + acceleration, Auto)
```

The `summary()` function produces a numerical summary of each variable
in a particular data set

```{r}
summary(Auto)
```

